# Regression Modeling and Analysis of Earthquakes  
## Kecheng Xu  
## March 19, 2015 

## Preliminaries

### Load libraries/data/create new variables
```{r loadData,results="hide"}
library(Hmisc)

data=read.csv('../data/all_month.csv',sep=",",header=TRUE)
dt=data[complete.cases(data),] ## remove the rows containing missing values
dt=subset(dt,dt$type=='earthquake')
dt$latCut <- cut2(dt$latitude, g = 5)
dt$lonCut <- cut2(dt$longitude, g = 5)
dt$nstCut <- cut2(dt$nst, g = 5)
dt$log10Depth <- log10(dt$depth-min(dt$depth)+1)
dt$time=as.character(dt$time)
day=substr(dt$time,1,10)
hour=substr(dt$time,12,19)
new.time=paste(day,hour)
dt$time=strptime(new.time,format="%Y-%m-%d %H:%M:%S")
summary(dt)
```
------

## Exploratory analysis

### Get minimum and maximum times and date downloaded (Methods/Data Collection)
```{r}
summary(dt$time)
```

### Look at patterns over time (Results paragraph 1)
```{r}
plot(dt$time, dt$mag, pch = 19,cex=0.5)
plot(dt$time, dt$depth, pch = 19,cex=0.5)
```

There does not appear to be a time trend in either variable. 


### Look at distribution of magnitudes (Results paragraph 2)
```{r}
mean(dt$mag < 3)
mean(dt$mag > 3 & dt$mag < 5)
```
Most earthquakes are small (< 3) or medium (>3 and < 5)

### Look at distribution of depths (Results paragraph 2)

```{r}
hist(dt$depth, col = "grey",main="Histgram of the Distribution of Depths",
     xlab="depth",ylab="frequency")
hist(dt$log10Depth, col = "grey",main="Histgram of the Distribution of log 10 based Depths",
     xlab="depth",ylab="frequency")
```

-------

## Modeling 

### Fit a model with no adjustment (results - paragraph 3)

```{r}
# Fit model with no adjustment variable
lm.no.adjust <- lm(dt$mag ~ dt$log10Depth)

# Plot residuals, colored by different variables (latitude, longitude, number of sites observing the quake)
par(mfrow = c(1, 3))
plot(dt$log10Depth, lm.no.adjust$residuals, col = dt$latCut, pch = 19)
plot(dt$log10Depth, lm.no.adjust$residuals, col = dt$lonCut, pch = 19)
plot(dt$log10Depth, lm.no.adjust$residuals, col = dt$nstCut, pch = 19)
```
It appears there are some non-random patterns here.


### Now fit a model with factor adjustment for latitude, longitude, and number of sites (results - paragraph 3)

```{r lmFinalChunk, dependson="loadData"}
lm.final <- lm(dt$mag ~ dt$log10Depth + dt$latCut + dt$lonCut + dt$nst)

par(mfrow = c(1, 3))
plot(dt$log10Depth, lm.final$residuals, col = dt$latCut, pch = 19)
plot(dt$log10Depth, lm.final$residuals, col = dt$lonCut, pch = 19)
plot(dt$log10Depth, lm.final$residuals, col = dt$nstCut, pch = 19)
```

## Get the estimates and confidence intervals

```{r, dependson="lmFinalChunk"}
## The estimate from summary
summary(lm.final)

## The confidence interval from confint
confint(lm.final)
```











